Redshift Evolution of the Type Ia Supernova Population Stretch Distribution

### 2020-01-27 Nicolas.tex
Big update. Version 1.2.

\begin{document}
\title{Redshift evolution of the SN stretch distribution}

@@

\author{N. Nicolas\inst{1}
    \and M. Rigault\inst{2}
    \and R. Graziani\inst{2}     
    \and M. Briday\inst{1}
    \and Y. Copin\inst{1}    
    \and Y. Kim\inst{1}
}

@@

\abstract{Type Ia supernovae (SNe Ia) allow for the construction of the Hubble
    diagram, giving us information about the Universe's expansion and its
    fundamental components, one of which is dark energy. But systematic
    uncertainties are now starting to be limiting in our ability to measure
those parameters. In particular, the physics of SNe Ia is still mostly unknown,
and is thought not to change in time/with the redshift.}
    {In an attempt to reduce those uncertainties, we try to find an empirical
    law describing SNe Ia's length of explosion (stretch) evolution with the
redshift.}
    {We started by getting a complete sample representing all of the stretch
        distribution that Nature can give us, before using LsSFR measurements,
        an age tracer which evolution with redshift is known, that has been
        shown to have a strong correlation with the stretch. We compare their
        AICc, an estimator of the relative quality of statistical models that
    includes the number of free parameters, to determine which ones describe
best the data.}
    {Models with an evolution of the stretch with the redshift have a better
    AICc than the ones without.}
    {We find that implementing these models allows us to fit the data better
    than models without stretch evolution.}

@@

Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
have enabled the discovery of the acceleration of the Universe's expansion
\citep{riess1998, perlmutter1999}. They remain today a key cosmological probe to
understand the properties of dark energy (DE) as it is the only tool able to
precisely map the recent expansion rate $z<0.5$, when DE is driving it
\citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
Hubble Constant (H$_0$) provided one can calibrate their absolute magnitude
\citep{riess2016, freedman2019}. Interestingly, the value of H$_0$ derived when
the SNeIa are anchored on Cepheids \citep[the SH0ES
project][]{riess2009,riess2016} is $\sim5~\sigma$ higher than what is predicted
by cosmic microwave background (CMB) data measured by Planck assuming the
standard $\L$CDM or when the SN luminosity is anchored at intermediate redshift
by the baryon acoustic oscillation (BAO) scale \citep{riess2019, reid2019,
planck2016, feeney} \nn{changed ried to reid; changed planck to planck2016;
feeney = https://tinyurl.com/vg6uufm ?}. While using the tip of the red giant
branch technique in place of the Cepheids seem to favor lower values of H$_0$
\citep{freedman2019, recalibrationpaper} \nn{changed "freedman" to
"freedman2019" ; recalibrationpaper = ?} time delay measurements from strong
lensing seem to also favor high H$_0$ values \citep{wong2019}.

The H$_0$ tension has received a lot of attention as it could be a sign of new
fundamental physics. Yet, no simple solution is able to accommodate this H$_0$
tension when accounting for all other probes and the current most promising
scenario appears to be a burst of expansion at the matter-radiation decoupling
moment caused by a (fine tuned) early dark energy \citep{poulin2019} \nn{changed
"poulin" to "poulin 2019"}.

Alternatively, systematic effects affecting one or several of the aforementioned
analysis could also explain the tension. In \cite{rigault2015} we suggested that
SNe Ia from the Cepheid calibrator sample differ significantly to the one from
the Hubble flow sample that enables to derive H$_0$. Indeed, Cepheid are very
young stars and the existence of Cepheids in a galaxy implies that the host is
star forming. In \citep{rigault2015} we claimed that a more than 90\% of the
calibrator SNe Ia arise from a young-progenitor population while this population
only accounts for about half of the Hubble flow sample. Following
\cite{sullivan2010} and \citep{rigault2013} \nn{should use cite or citep?}, we
further showed that young and old progenitor SNe Ia have a different magnitude
of about 0.15 mag. This has been confirmed for modern standardization technique
\citep[SALT2.4][]{guy2010,betoule2014} with high significance ($\sim 6\sigma$)
in \citep{rigault2018} as well as independently by \cite{roman2019} \nn{not
found on ADS: https://tinyurl.com/uemattg} at $\sim 7\sigma$. As detailed in
\citep{rigault2015}, if the magnitude difference between young and old SNe Ia is
indeed of $0.15$ mag and if SNe Ia from the calibrator sample is fully dominated
by young SNeIa, the effect on H$_0$ of the existence of these two SNe Ia
populations is of 3.5\%. Accounting for the fact that another environmental
effect (the mass-step) is already included in the SH0ES analysis, the net bias
in the SH0ES analysis is expected closer to 3\% \citep[see table 6
of][]{rigault2015}.

SHOULD BE SHORTENED A LOT: However, the importance of this astrophysical bias in
the direct measurement of H$_0$ is still highly debated. We also highlight that
it could not explain the high value of H$_0$ derived by strong lensing. First,
\cite{jones2015} \nn{put the ref but nos sure it's the correct one} did not find
any environmental bias when extending the \citep{rigault2015} study. It was more
recently reported in \citep{jones2019} \nn{put the ref but nos sure it's the
correct one} that indeed local host analysis has significant influence on the SN
magnitude but much more importantly than claimed in \citep{rigault2018} and with
a connection with the progenitor physics and/or a line of sight effect that is
unclear. More importantly, \cite{riess2016} attended to mimic the Cepheid sample
selection function on the Hubble flow sample by removing SNe Ia from the latter
if they classified their host as early-type or non-star forming. They measured
that doing so has no effect on the measurement of H$_0$.

This paper is part of a series of papers where we test the validity of the two
SNe Ia population models developed in \citep{rigault2018}.Briday et al. in prep
will present how using different environmental tracer is affecting our ability
to distinguish the two populations, and Rigault et al. in prep will further
study the connection with H$_0$. Here we analyze predictions concerning the SNe
Ia redshift drifts: (1) that the distribution of the SNIa stretch, a purely
intrinsic SNIa property, is evolving as a function of redshift and (2) that this
population-drift follows the relation that \citep{rigault2018} derived assuming
that one population is “young“ with a rate proportional to the star formation
rate in the Universe and the second is "old" with a rate proportional to the
stellar mass.

The concept of the SNeIa age dichotomy arised when the SNIa rate has first been
studied. \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have
shown that the relative SNe Ia rate in galaxies could only be explained if two
populations existed, one young, following the host star formation activity, and
one old following the host stellar mass (the so called “prompt and delayed“ or
“A+B“ model). In \cite{rigault2018} we used the specific star formation rate at
the SN location (Local sSFR or LsSFR) to classify which are the prompts (those
with a high LsSFR) and which are the delayed (those with low LsSFR). Since the
first SNe Ia host analysis, the SN stretch has been known to be strongly
correlated with the SN host properties \citep{hamuy1996, hamuy2000} \nn{put
hamuy1996 as most cited of https://tinyurl.com/sjlbdz8} and it has been
extensively confirmed since \citep[e.g.][]{neill2009, sullivan2010,
lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013, rigault2013,
pan2014} \nn{added all the refs}. Following the “A+B“ model and the connection
between SN stretch and host properties, \citep{howell2009} \nn{changed
howell20xx to howell2009
(https://ui.adsabs.harvard.edu/abs/2009ApJ...691..661H/abstract)} first
discussed the potential redshift drift of the SN stretch distribution. In this
paper we revisit this analysis with the most recent SNe Ia dataset and we test
the validity of the two SNIa age populations model developed by
\citep{rigault2018}.

We present in section~\ref{sec:sample} the sample we are using for this
analysis, which is based on the pantheon dataset \citep{scolnic2018a}. We
discuss the importance of obtaining a "complete" sample, i.e. representative of
the true underlying SNe Ia distribution, and how we build one from the Pantheon
sample.  We then present in section~\ref{sec:modeling} our modeling of the
distribution of stretch as a function of redshift based on the relative stretch
distribution of young and old SNeIa. Our results are presented in
section~\ref{sec:results} where we test if the distribution of stretch does
evolve as a function of redshift and if the age-model is in good agreement with
this evolution. We discuss our results and conclude in section~\ref{sec:ccl}.

Throughout the paper we use the Planck 2015 cosmology from the astropy library.

\section{A "complete" Sample}
\label{sec:sample}

We base our analysis on the latest SNe Ia compilation: the Pantheon catalog from
\cite{scolnic2018a}.  \mr{TO BE REPHRASED} Using this dataset, we aim at probing
if the average stretch, an intrinsic SN Ia property, evolves as a function of
redshift or not; the underlying question being that of the SNIa population
redshift drift.  The stretch has been shown to depend on the SN environment star
formation activity
\citep[e.g.][]{howell2007,sullivan2010,childress2013,rigault2013,rigault2018}
and because the star formation rate is about 40 times larger at the high
redshift end of the pantheon sample than at its lower redshifts
\citep[e.g.,][and references therein]{tasca2015}, one might think that simply
comparing the redshifts distribution in a few bins of redshift could be
sufficient to probe any SN stretch redshift drift.  In practice, selection
effects are affecting the observed SN stretch distributions, which could
significantly differ from the true underlying one if these selection effects
become significant. Indeed, because the observed SNIa magnitude correlates with
the lightcurve stretches (and colors) the first SNIa that such a survey will
miss will the reddest and the lowest-stretch ones. Consequently, if selection
effects is not accounted for, one might confuse true population drift with
selection effect and conversely.

In a magnitude limited survey, providing sufficient (and random) spectroscopic
follow-up for acquiring typing and host redshift, the selection effects should
be negligible bellow a given redshift where all SNe Ia, even the faintest, could
be observed.  In contrast, targeted surveys have highly complex selection
functions and we will discard them from our analysis. Fortunately, modern SN
cosmology sample such has the pantheon dataset are now dominated by magnitude
limited survey.

We present fig.~\ref{fig:maglim} the lightcurve stretch and color distributions
of SNeIa from these surveys, namely: PanStarrs \citep[PS1][]{ref}, the Sloan
Digital Sky Survey \citep[SDSS][]{ref} and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}.  An ellipse in the salt2.4 $(x_1, c)$ plan with
major axis $x_1=3$ and manor axis $c=0.3$ encapsulates the full distribution
(see also \citealt{bazin2011} and \citealt{campbell2013} for a similar contour
but using a less conservative $c=0.2$ for the latter). Assuming the SN absolute
magnitude with $x_1=0$ and  $c=0$ is $M=-19.36$ \citep{kessler2009,scolnic2014}
we can derive the absolute standardized magnitude at maximum of light along the
aforementioned ellipse given the standardization coefficient $\alpha=0.16$ and
$\beta=3.14$ from \cite{scolnic2018a}.  The faintest SNIa is that with
$x_1=-1.66$ and $c=0.25$ and has an absolute standardized magnitude at peak in
Bessel-b band of $-18.31\,\mathrm{mag}$. Since we need to see this object
typically a week before and 10 days after peak to build a suitable lightcurve,
the effective absolute limiting magnitude therefore is
$\sim-18.00\,\mathrm{mag}$.  Hence, given the $5\sigma$ point source detection
magnitude limit of a magnitude limited survey, one can derive the maximum
redshift above which the faintest SNeIa will start to be missed.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/zmax_maglim_all.pdf}
    \caption{SALT2.4 stretch and color lightcurve parameters of SNeIa from the
        SDSS, PS1 and SNLS samples from the pantheon catalog. The individual SN
        data are shown as blue dots and a 2D histogram is shown in gray to
        highlight point density. The ellipse encapsulating all the SN data is
        displayed colored by the effective standardized absolute magnitude using
    the $\alpha$ and $\beta$ standardization coefficients from
\citep{scolnic2018a}.}
    \label{fig:maglim}
\end{figure}

SNLS typically acquires SNeIa in the redshift range $0.4<z<0.8$. At these
redshifts the rest-frame Bessel-b band roughly corresponds to the SNLS-$i$
filter that has a 24.8 mag $5\sigma$
depth\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{lim}=0.60$, in perfect
agreement with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}.
Similarly, PS1 observes SNeIa in the range $0.2<z<0.4$, their $g$-band $5\sigma$
depth is 23.1 mag \citep{rest2014}, which yields to $z_{lim}=0.30$.  Figure~6 of
\citep{scolnic2018a} suggests a more conservative $z_{lim}$ of 0.27 for the PS1
catalog.  In similar redshift range, SDSS has a limiting magnitude of 22.5
\citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However, the
SDSS survey were more sensitive to limited spectroscopic resources.
\cite{kessler2009} present that during year-1 of SDSS, SNeIa with $r-mag<20.5$
were favored for spectroscopic follow up, corresponding to the redshift cut at
$0.15$. For the rest of the SDSS survey, additional spectroscopic resources were
used, such that \cite{kessler2009} and \cite{dilday2008} show a relative
completeness up to $z_{lim}=0.2$.  Following these analysis we will use
$z_{lim}=0.2$ as the baseline SDSS redshift limit for the rest of the analysis.
We could have chosen more conservative redshift limits for each of these
surveys, namely $z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS and
$z_{lim}=0.55$ for SNLS, following Fig.~14 of \citealt{perrett2010}, and we will
also present results obtained when doing so.  We present the redshift
distribution of these three surveys in fig.~\ref{fig:cuts}, where we see that
the redshift limit roughly correspond to the pick of these histograms, as
expected.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/hist_surveys_cuts.pdf}
    \caption{Histograms of the 3 surveys on which we applied cuts to get rid of
    selection effects, as discussed in section \ref{sec:sample}. The SNe Ia
between the conservative and non-conservative cuts are shown in transparent.}
    \label{fig:cuts}
\end{figure}

In addition, we use the SNeIa from the nearby supernovae factory
\citep[SNfactory][]{aldering2004} \nn{added ref assuming
https://ui.adsabs.harvard.edu/abs/2004APS..APR.V4003A/abstract} published in
\citep{rigault2018} and that have been discoverd from no-targeted survey (114
SNeIa, see their section~3 and 4.2.2). As the SNe search were much deeper than
the spectrophotometric follow-up, SNfactory SNe Ia within a redshift range of
$0.02<z<0.09$ should also be a random sampling of the underlying SN population.
Data from \citep{rigault2018} are within these redshift boundaries. The HST
sample from Pantheon follows the same logic and we therefore kept it entirely
\citep{FIND BACK THE REF} \nn{lots of choices: https://tinyurl.com/rcsx7cw}.

\begin{table}
    \centering
    \caption{Source surveys of SNeIa used in our "complete" sample. Conservative
    cuts are indicated in brackets.}
    \label{tab:sample}
    \begin{tabular}{l c c}
    \hline\hline\\[-0.8em]
        Survey & N$_{\mathrm{SN}}$  & z$_{max}$ \\[0.15em]
        \hline\\[-0.8em]
        SNf & 114 & --\\[0.30em]
        SDSS & 167(82) & 0.20(0.15)\\[0.30em]
        PS1 & 160(122) & 0.30(0.27) \\[0.30em]        
        SNLS & 102(78) & 0.60(0.55)\\[0.30em]
        HST & 26 & --\\[0.30em]
        \hline
    \end{tabular}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%
%       INFO          %
%%%%%%%%%%%%%%%%%%%%%%%
%Fraction of random SDSS stretchs being < mean_conservative =  0.299
%Fraction of random PS1 stretchs being < mean_conservative =  0.254
%Fraction of random SNLS stretchs being < mean_conservative =  0.417

@@

In \cite{rigault2018} we presented a modeling of the evolution of the fraction
of prompt and delayed SNeIa as a function of redshift following former work on
rates and delay time distributions; see e.g. the prompt and delayed SNeIa model,
aka the "A+B" model,
\cite{mannucci2005,scannapieco2005,sullivan2006,aubourg2008} and
\cite{maozmannucci2014} for a review on delay time distributions of SNeIa. In
short, we assume that the number of prompt SNeIa follows the star formation
activity in the Universe while the number of delayed SNeIa follows the number of
Gyr-old stars in the Universe, i.e. the stellar mass. Hence, if we denote
calling $\delta(z)$ (resp. $\phi(z)$) the fraction of young (resp. old) SNeIa in
the Universe as a function of redshift, then their ratio is expected to follow
the evolution of the specific star formation rate which goes as $(1+z)^{2.8}$
until $z\sim2$ \citep{tasca2015}. Hence since there is  $\sim50\%$ of young
SNeIa at $z\sim0.05$ \citep{rigault2013,rigault2018}, in agreement with rate
expectations \citep{mannucci2006,rodney2014}, then, we find in
\cite{rigault2018}:
\begin{align}
    \label{eq:delta}
    \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}\,
                  \mathrm{or,}\\
    \psi(z) & = \left( K \times (1+z)^{+2.8} +1 \right)^{-1},
\end{align}

where $K=0.87$ given the fraction of young in the SNfactory sample. This
modelisation is comparable to the evolution predicted by \cite{childress2014}
based on SN rates in galaxies depending on their quenching time as a function of
their stellar mass.

\subsection{With LsSFR data}
\label{sec:modelpy}

>>>>

\begin{document}
%\title{Evolution of the Type Ia Supernova Underlying Stretch Distribution as a Function of Redshift}
\title{Redshift Evolution of the Type Ia Supernova Population Stretch Distribution}

@@

\author{N. Nicolas
\thanks{n.nicolas@ipnl.in2p3.fr, equal contribution}
\inst{1}
    \and M. Rigault
\thanks{m.rigault@ipnl.in2p3.fr, equal contribution}
\inst{2}
    \and R. Graziani\inst{2}     
    \and M. Briday\inst{1}
    \and Y. Copin\inst{1}    
    \and Y. Kim\inst{1}
}

@@

\abstract{\mr{ABSTRACT NOT DONE YET.}
Type Ia supernovae (SNe Ia) allow for the construction of the Hubble
    diagram, giving us information about the Universe's expansion and its
    fundamental components, one of which is dark energy. But systematic
    uncertainties are now starting to be limiting in our ability to measure
    those parameters. In particular, the physics of SNe Ia is still mostly
    unknown, and is thought not to change in time/with the redshift.}
    {In an attempt to reduce those uncertainties, we try to find an empirical
    law describing SNe Ia's length of explosion (stretch) evolution with the
redshift.}
    {We started by getting a complete sample representing all of the stretch
        distribution that Nature can give us, before using LsSFR measurements, an
        age tracer which evolution with redshift is known, that has been shown
        to have a strong correlation with the stretch. We compare their AIC, an
    estimator of the relative quality of statistical models that includes the
number of free parameters, to determine which ones describe best the data.}
    {Models with an evolution of the stretch with the redshift have a better
    AIC than the ones without.}
    {We find that implementing these models allows us to fit the data better
    than models without stretch evolution.}

@@

Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
have enabled the discovery of the acceleration of the Universe's expansion
\citep{riess1998, perlmutter1999}. They remain today a key cosmological probe to
understand the properties of dark energy (DE) as it is the only tool able to
precisely map the recent expansion rate $z<0.5$, when DE is driving it
\citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
Hubble Constant (H$_0$), provided one can calibrate their absolute magnitude
\citep{riess2016, freedman2019}. Interestingly, the value of H$_0$ derived when
the SNe~Ia are anchored on Cepheids \citep[the SH0ES
project][]{riess2009,riess2016} is $\sim5\sigma$ higher than what is predicted
by cosmic microwave background (CMB) data measured by Planck assuming the
standard $\L$CDM\nn{,} or when the SN luminosity is anchored at intermediate
redshift by the baryon acoustic oscillation (BAO) scale
\citep{riess2019,reid2019,planck2016, feeney2019}.  While using the tip of the
red giant branch technique in place of the Cepheids seem to favor an
\mr{intermediate} value of H$_0$ \citep{freedman2019}, time delay measurements
from strong lensing seem to also favor high H$_0$ values \citep{wong2019}.

The H$_0$ tension has received a lot of attention as it could be a sign of new
fundamental physics. Yet, no simple solution is able to accommodate this H$_0$
tension when accounting for all other probes \citep{knox2019} and the current
most promising scenario appears to be a burst of expansion at the
matter-radiation decoupling moment caused by a (fine tuned) early dark energy
\citep{poulin2019}.

Alternatively, systematic effects affecting one or several of the aforementioned
analysis could also explain the tension.  In \cite{rigault2015} we suggested
that SNe Ia from the Cepheid calibrator sample differs by construction from  the
Hubble flow as it strongly favors young stellar population environments where
one could find Cepheids. This selection effect would impact the derivation of
H$_0$ if SNe~Ia from young and older environments differ in average magnitudes.

For the last decades, numerous analysis have studied the relation between SNe~Ia
and host properties\nn{,} finding first that the standardized SNe~Ia magnitude
significantly depends on the host stellar-mass, SNe~Ia from high-mass host
\nn{being brighter on average}
\cite[e.g.][]{kelly2010,sullivan2010,childress2013,betoule2014,rigault2018}.
This mass-step correction is currently used in cosmological analyses
\citep[e.g.][]{betoule2014,scolnic2018a}, including for deriving H$_0$
\citep{riess2016,riess2019}. Yet, the underlying connection between the SN and
the host remains unclear when using global properties such as the host stellar
mass, which raises the question of the accuracy of the correction. More
recently, studies have used the local SN environment to probe more direct
connections between the SN and its environments \citep{rigault2013}\nn{,}
showing that local age-tracers such as the local specific star formation rate or
the local color are more strongly correlated with the standardized SN magnitude
\citep{rigault2018,roman2018}, suggesting the age to be the driving parameter
underlying the mass-step. If true, this would have a significant impact for
cosmology because the redshift drift magnitude correction to apply might
strongly vary \citep{rigault2013,childress2014,scolnic2018a}.  Yet, the
importance of local SN environmental study remains highly debated
\cite[e.g.][]{jones2015,jones2019} and especially the impact it has on the
derivation of H$_0$ \cite{jones2015,riess2016,riess2018,rose2019}.

In this paper, we take a step aside to probe the validity of our modeling of the
SN population that we claim to be constituted of two age-populations
\citep{rigault2013,rigault2015,rigault2018}: one old and one young, the former
having on average lower lightcurve stretches and being brighter after
standardization. We indeed use the correlation between the SN age probed with
local specific star formation rate in \cite{rigault2018} and the SN stretch to
model the expected evolution of the underlying SN stretch distribution as a
function of redshift. This modeling relies on three assumptions: (1) there
\nn{are indeed} 2 populations of SNe~Ia; (2) the relative fraction of each of
these populations as a function of redshift follows the age-modeling made in
\citep{rigault2018} and (3) the underlying distribution of stretch for each
age-sample can be modeled and do not significantly drift. This paper aims at
testing this with data from the literature.

% SERIES OF PAPER
%This paper is part of a series of papers where we test the validity of the two
%SNe Ia population models developed in \citep{rigault2018}. Briday et al. in
%prep will present how using different environmental tracer is affecting our
%ability to distinguish the two populations, and Rigault et al. in prep will
%further study the connection with H$_0$. Here we analyze predictions concerning
%the SNe Ia redshift drifts: (1) that the distribution of the SN~Ia stretch, a
%purely intrinsic SN~Ia property, is evolving as a function of redshift and (2)
%that this population-drift follows the relation that \citep{rigault2018}
%derived assuming that one population is “young“ with a rate proportional to the
%star formation rate in the Universe and the second is "old" with a rate
%proportional to the stellar mass.

The concept of the SNe~Ia age dichotomy arised when the SN~Ia rate has first
been studied. \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008}
have shown that the relative SNe~Ia rate in galaxies could only be explained if
two populations existed, one young, following the host star formation activity,
and one old following the host stellar mass (the so called “prompt and delayed“
or “A+B“ model). In \cite{rigault2018} we used the specific star formation rate
at the SN location (Local sSFR or LsSFR) to classify which are the prompts
(those with a high LsSFR) and which are the delayed (those with low LsSFR).
Furthermore, since the first SNe~Ia host analysis, the SN stretch has been known
to be strongly correlated with the SN host properties \citep{hamuy1996,
hamuy2000}, which has been extensively confirmed since \citep[e.g.][]{neill2009,
sullivan2010, lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013,
rigault2013, pan2014}. Following the “A+B“ model and the connection between SN
stretch and host properties, \cite{howell2007} first discussed the potential
redshift drift of the SN stretch distribution. In this paper we revisit this
analysis with the most recent SNe~Ia dataset..

We present in section~\ref{sec:sample} the sample we are using for this
analysis, which is based on the Pantheon dataset \citep{scolnic2018a}. We
discuss the importance of obtaining a "complete" sample, i.e. representative of
the true underlying SNe Ia distribution, and how we build one from the Pantheon
sample.  We then present in section~\ref{sec:modeling} our modeling of the
distribution of stretch as a function of redshift based stretch distribution of
young and old SNe~Ia. Our results are presented in section~\ref{sec:results}
where we test if the SN stretch does evolve as a function of redshift and if the
age model is in good agreement with this evolution. Consequence for cosmology of
our results are briefly discussed in section~\ref{sec:bbc} and we conclude in
section~\ref{sec:ccl}.

%Throughout the paper we use the Planck 2015 cosmology from the astropy library.

\section{A ''complete'' Sample}
\label{sec:sample}

We base our analysis on the latest SNe~Ia compilation: the Pantheon catalog from
\cite{scolnic2018a}.  A naive approach to test the SN stretch redshift drift
would be to simply compare the observed SN stretch distributions in a few bins
of redshift.  In practice, selection effects are affecting the observed SN
stretch distributions. Indeed, because the observed SN~Ia magnitude correlates
with the lightcurve stretch (and color), the first SN~Ia that a
magnitude-limited survey will miss are the reddest and the lowest-stretch ones.
Consequently, if selection effects are not accounted for, one might confuse true
population drift with selection effects, and conversely.

\nn{Assuming} sufficient (and random) spectroscopic follow-up for acquiring
typing and host redshift, the selection effects of magnitude-limited survey
should be negligible bellow a given redshift where even the faintest \nn{SN~Ia}
could be observed.  In contrast, targeted surveys have highly complex selection
functions and will be discarded from our analysis. Fortunately, modern SN
cosmology samples such as the pantheon dataset are now dominated by magnitude
limited surveys.

We present fig.~\ref{fig:maglim} the lightcurve stretch and color distributions
of SNe~Ia from these surveys, namely: PanStarrs \citep[PS1][]{rest2014}, the
Sloan Digital Sky Survey \citep[SDSS][]{frieman2008} and the SuperNovae Legacy
Survey \citep[SNLS][]{astier2006}. An ellipse in the salt2.4 $(x_1, c)$ plan
with semi-major axis $x_1=3$ and semi-minor axis $c=0.3$ encapsulates the full
distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011} and
\citet{campbell2013} for similar contours, the second using a more conservative
$c=0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and  $c=0$ is
$M_0=-19.36$ \citep{kessler2009,scolnic2014} we can derive the absolute
standardized magnitude at maximum of light along the aforementioned ellipse
given the standardization coefficient $\alpha=0.156$ and $\beta=3.14$ from
\cite{scolnic2018a}. The faintest SN~Ia is that with $x_1=-1.66$ and $c=0.25$
and has an absolute standardized magnitude at peak in Bessel-b band of
$-18.31\,\mathrm{mag}$. Since one need to detect this object typically a week
before and 10 days after peak to build a suitable lightcurve, the limiting
\nn{standardized} absolute magnitude effectively is approximately
$-18.00\,\mathrm{mag}$. Hence, given the $5\sigma$ point source detection
magnitude limit of a magnitude limited survey $m_{lim}$, one can derive the
maximum redshift above which the faintest SNe~Ia will be missed \nn{using the
link between apparent magnitude, redshift and absolute magnitude given by:}
\begin{equation}\label{eq:m_lim}
    \nn{m = \mu(z) + \underbrace{M_{min}^{t_0-10}(x_1, c)}_{=-18.00\,\text{mag}}
    < m_{lim}}
\end{equation}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_all.pdf}
    \caption{SALT2.4 stretch and color lightcurve parameters of SNe~Ia from the
        SDSS, PS1 and SNLS samples from the pantheon catalog. The individual SN
        data are shown as blue dots and a 2D histogram is shown in gray to
        highlight point density.  The ellipse ($x_1=\pm3\,\mathrm{;}\,c=\pm0.3$)
    is displayed, colored by the corresponding standardized absolute magnitude
using the $\alpha$ and $\beta$ coefficients from \citep{scolnic2018a}.}
    \label{fig:maglim}
\end{figure}

SNLS typically acquires SNe~Ia in the redshift range $0.4<z<0.8$. At these
redshifts the rest-frame Bessel-B band roughly corresponds to the SNLS-$i$
filter that has a 24.8 mag $5\sigma$
depth\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{lim}=0.60$, in perfect
agreement with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}.
Similarly, PS1 observes SNe~Ia in the range $0.2<z<0.4$, their $g$-band
$5\sigma$ depth is 23.1 mag \citep{rest2014}, which yields to $z_{lim}=0.30$ in
agreement with, e.g., Figure~6 of \cite{scolnic2018a}. This figure could also
suggest a more conservative $z_{lim}$ of 0.27; both will be considered as
discussed below.  In similar redshift range, SDSS has a limiting magnitude of
22.5 \citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However,
the SDSS surveys were more sensitive to limited spectroscopic resources.
\cite{kessler2009} \nn{pointed out} that during year-1 of SDSS, SNe~Ia with
$r-mag<20.5$ were favored for spectroscopic follow up, corresponding to the
redshift cut at $0.15$. For the rest of the SDSS survey, additional
spectroscopic resources were used, such that \cite{kessler2009} and
\cite{dilday2008} show a relative completeness up to $z_{lim}=0.2$.  Following
these analyses we will use $z_{lim}=0.2$ as the baseline SDSS redshift limit for
the rest of \nn{this study}.  For the rest of the analysis, we will consider
both cases, the \nn{aforementioned} redshift limits as well as more conservative
cuts for each of these surveys, namely $z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$
for PS1 and $z_{lim}=0.55$ for SNLS, following Fig.~14 of \citealt{perrett2010}.

The complete sample selection is summarized table~\ref{tab:sample} and the
redshift distribution of these three surveys in are shown fig.~\ref{fig:cuts}.
We see in this figure that the redshift limit we have selected roughly
correspond to the peak of these histograms as expected.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{Article_figures/hist_surveys_cuts.pdf}
    \caption{From top to bottom: histogram of SN~Ia redshifts for SDSS, PS1 and
    SNLS from the pantheon dataset, respectively. The colored part of the
distribution shows which SNe~Ia has been kept in our analysis for they are
supposedly free from selection bias (see section~\ref{sec:sample}). The
full(light) color responds to the conservative(nominal) cut.}
    \label{fig:cuts}
\end{figure}

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory][]{aldering2004} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their section~3 and
4.2.2). As the searches were much deeper than the spectrophotometric follow-up
made by SNfactory \nn{(SNf)}, SNe~Ia within a redshift range of $0.02<z<0.09$
\citep[as in ][]{rigault2018} should also be a random sampling of the underlying
SN population. The SNf data is particularly useful for studying SN property
drift as it enable us to have a large SN~Ia sample at $z<0.1$.  The HST sample
from Pantheon follows the same logic of having a search deeper than the
follow-up and we therefore kept it entirely \citep{FIND BACK THE REF}, see
table~\ref{tab:sample}.

\begin{table}
    \centering
    \caption{\nn{Source surveys and number of SNe~Ia used in our "complete"
    sample. Conservative cuts are indicated in brackets.}}
    \label{tab:sample}
    \begin{tabular}{l c c}
    \hline\hline\\[-0.8em]
        Survey & z$_{lim}$ & N$_{\mathrm{SN}}$ \\[0.15em]
        \hline\\[-0.8em]
        SNf & -- & 114\\[0.30em]
        SDSS & 0.20(0.15) & 167(82)\\[0.30em]
        PS1 & 0.30(0.27) & 160(122)\\[0.30em]        
        SNLS & 0.60(0.55) & 102(78)\\[0.30em]
        HST & -- & 26\\[0.30em]
        Total & -- & 569(422)\\[0.30em]
        \hline
    \end{tabular}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%
%       INFO          %
%%%%%%%%%%%%%%%%%%%%%%%
%Fraction of random SDSS stretches being < mean_conservative =  0.299
%Fraction of random PS1 stretches being < mean_conservative =  0.254
%Fraction of random SNLS stretches being < mean_conservative =  0.417

@@

In \cite{rigault2018} we presented a modeling of the evolution of the fraction
of prompt and delayed SNe~Ia as a function of redshift following former work on
rates and delay time distributions \citep[e.g.,][]{mannucci2005,
scannapieco2005, sullivan2006, aubourg2008, childress2014, maozmannucci2014}.
In short, we assumed that the number of prompt SNe~Ia follows the star formation
activity in the Universe while the number of delayed SNe~Ia follows the number
of Gyr-old stars in the Universe, i.e. the stellar mass. Hence, if we denote
$\delta(z)$ (resp. $\psi(z)$) the fraction of young (resp. old) SNe~Ia in the
Universe as a function of redshift, then their ratio $\delta/\psi$ is expected
to follow the evolution of the specific star formation rate which goes as
$(1+z)^{2.8}$ until $z\sim2$ \citep[e.g.,][]{tasca2015}. Hence since
\nn{$\delta(0.05) \sim \psi(0.05)$} \citep{rigault2013,rigault2018,wiseman2020},
in agreement with rate expectations \citep{mannucci2006,rodney2014}, \nn{and
$\delta + \psi = 1$,} then we find in \cite{rigault2018}:
\begin{align}
    \label{eq:delta}
    \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}\,
    \mathrm{and}\\
    \psi(z) & = \left( K \times (1+z)^{+2.8} +1 \right)^{-1}
\end{align}
where $K=0.87$. This modelisation is comparable to the evolution predicted by
\cite{childress2014} based on SN rates in galaxies depending on their quenching
time as a function of their stellar mass.

\subsection{\nn{Base underlying distribution}}
\label{sec:basemodel}
